### 2021-11-13 NSF2021/Description/priorsupport.tex

\noindent
{\bf Yuan-Nan Young}: {\em NSF-DMS-1222550, Mathematical and
experimental study of lipid bilayer shape and dynamics mediated by
surfactants and proteins}, \$212,603, 9/15/2012 - 08/31/2016 (with
no-cost extension), PI. {\em Intellectual merit:} The focus of this
grant is modeling the interaction between a pure lipid bilayer membrane
with surfactant, cholesterol and protein.
%Hence, there is no overlap between these grants and the current proposal.

@@

\noindent
{\it Broader impacts:} 
One PhD student (Szu-Pei Fu) was funded to work with YNY, and work has
resulted in seven papers
\cite{Nganguia2013_PoF,Nganguia2013_PRE,Young2014_JFM,Young2015_PoF,Nganguia2015_CiCP,Pak2015_PNAS,fu2015pre}.
YNY has been actively involved with promotion of underrepresented
students at NJIT. The other PhD student (Herve Nganguia) is African.
YNY has taught a broad spectrum of courses in fluid mechanics and
applied math modeling.

@@

\noindent
{\it Broader impacts:} 
A second-year PhD student (Jake Cherry) has been assigned to this
project. Since the funding just began, results have not yet
been published or presented.

@@

\begin{wrapfigure}[8]{r}{0.56\textwidth}
\vspace{-19pt}
\definecolor{barblue}{RGB}{51,102,254}
%\renewcommand\sfdefault{phv}
%\renewcommand\mddefault{mc}
%\renewcommand\bfdefault{bc}
%\sffamily
\begin{ganttchart}[
    canvas/.append style={fill=none, draw=black!5, line width=.75pt},
        x unit =4.5mm,
        y unit chart =\baselineskip,
    hgrid style/.style={draw=black!5, line width=.75pt},
    vgrid={*1{draw=black!5, line width=.75pt}},
    title/.style={draw=none, fill=none},
    title label font=\bfseries\footnotesize, % numbers across the top
    title label node/.append style={below=3pt},
    include title in canvas=false,
    bar label font=\mdseries\footnotesize\color{black!70}, 
    % titles across the left
    %bar label node/.append style={left=2cm},
    bar/.append style={draw=none, fill=barblue},
    bar incomplete/.append style={fill=barblue},
    bar progress label font=\mdseries\footnotesize\color{black!70},
    milestone label font=\mdseries\small\color{black!70},
        milestone left shift =0.9,
        milestone right shift =0.1,
        % Don't draw group bars
        group height =0,
        group peaks height =0,
        group label font =\bfseries\small,
]{1}{12}
\gantttitle[
    title label node/.append style={below left=3pt and -6pt}
]{QUARTERS:\quad1}{1}
\gantttitlelist{2,...,12}{1} \\
%\ganttgroup{Tasks\hfill}{1}{12}\\
  \ganttbar{Specific Aim 1 (2D)}{1}{4} \\
  \ganttbar{Specific Aim 1 (3D)}{8}{12}\\
\ganttbar{Specific Aim 2}{1}{8}\\
\ganttbar{Specific Aim 3}{1}{12}\\
\ganttbar{Program management}{1}{12}
\end{ganttchart}
  \vspace{-10pt}
\caption{\footnotesize Schedule for the proposed work, measured in
  quarters from the beginning of the project.}
\label{fig:schedule}
\end{wrapfigure}
\textbf{Project management}: 
%
The success of the proposed research requires complementary expertise
and collaborative efforts in physics, applied mathematics, algorithms,
and computing. Ryham has been working on mathematical modeling with a
strong analytical background. Young has been working on many areas of
computational fluid dynamics and applications to math biology for many
years. Quaife has been working on integral equation methods, fast
algorithms, and their applications to fluid dynamics for many years.
Their recent collaborative work on the HAP model in two dimensions has
provided a solid foundation for the proposed research.

\medskip

\textbf{Collaboration plan}: 
%
The management responsibility of this collaborative research will reside
with the lead PI (Ryham) for this endeavor. The research work is
structured to meet the tasks discussed in \S\ref{sec:proposed-work}.
%
The PIs, postdocs, and students will meet frequently on Zoom and
in person when possible. During the pandemic, the three PIs will meet
weekly on Zoom for a research meeting. Once the pandemic is under control, PI RR and PI YNY will conduct biweekly meetings in person, with PI BQ Zoom in from Florida.
The PIs will share software packages, paper
sources, and references on a common \textsf{Git} repository. The
resulting software packages will be posted on the \textsf{Github}
software repository.

\medskip

\textbf{Research Schedule:} The detailed schedule for the proposed work
is shown in Figure~\ref{fig:schedule}. Specific Aim 1 is split over two
periods to allow time for 3D code to be developed.

>>>>

\noindent
{\bf Yuan-Nan Young}: {\em NSF-DMS-1222550, Mathematical and
experimental study of lipid bilayer shape and dynamics mediated by
surfactants and proteins}, \$212,603, 9/15/2012 - 08/31/2016 (with
no-cost extension), PI. {\em Intellectual merit:} The focus of this
grant is modeling the interaction between a pure lipid bilayer membrane
with surfactant, cholesterol, and protein.
%Hence, there is no overlap between these grants and the current proposal.

@@

\noindent
{\it Broader impacts:} 
One PhD student (Szu-Pei Fu) was funded to work with YNY, and work has
resulted in seven papers~\cite{Nganguia2013_PoF, Nganguia2013_PRE,
Young2014_JFM, Young2015_PoF, Nganguia2015_CiCP,
Pak2015_PNAS,fu2015pre}. YNY has been actively involved with promotion
of underrepresented students at NJIT. The other PhD student (Herve
Nganguia) is African. YNY has taught a broad spectrum of courses in
fluid mechanics and applied math modeling.

@@

\noindent
{\it Broader impacts:} 
A third-year PhD student (Jake Cherry) has been assigned to this
project. One paper will be submitted before December
2021~\cite{lin-her-qua-che2021}.

@@

\begin{wrapfigure}[11]{r}{0.56\textwidth}
\vspace{-19pt}
%\definecolor{barblue}{RGB}{51,102,254}
%\renewcommand\sfdefault{phv}
%\renewcommand\mddefault{mc}
%\renewcommand\bfdefault{bc}
%\sffamily
\begin{ganttchart}[
    canvas/.append style={fill=none, draw=black!5, line width=.75pt},
        x unit =4.5mm,
        y unit chart =\baselineskip,
    hgrid style/.style={draw=black!5, line width=.75pt},
    vgrid={*1{draw=black!5, line width=.75pt}},
    title/.style={draw=none, fill=none},
    title label font=\bfseries\footnotesize, % numbers across the top
    title label node/.append style={below=3pt},
    include title in canvas=false,
    bar label font=\mdseries\footnotesize\color{black!70}, 
    % titles across the left
    %bar label node/.append style={left=2cm},
    bar/.append style={draw=none, fill=blue},
    bar incomplete/.append style={fill=blue},
    bar progress label font=\mdseries\footnotesize\color{black!70},
    milestone label font=\mdseries\small\color{black!70},
        milestone left shift =0.9,
        milestone right shift =0.1,
        % Don't draw group bars
        group height =0,
        group peaks height =0,
        group label font =\bfseries\small,
]{1}{12}
\gantttitle[
    title label node/.append style={below left=3pt and -6pt}
]{QUARTERS:\quad1}{1}
\gantttitlelist{2,...,12}{1} \\
%\ganttgroup{Tasks\hfill}{1}{12}\\
\ganttbar[bar/.append style={fill=red}]{Specific Aim 1}{1}{12} \\
\ganttbar[bar/.append style={fill=blue}]{Specific Aim 2}{1}{12}\\
\ganttbar[bar/.append style={fill=green}]{Specific Aim 3}{1}{12}\\
\ganttbar[bar/.append style={fill=black}]{Program Management}{1}{12}
\ganttbar[bar/.append style={fill=magenta}]{}{4}{4}
\ganttbar[bar/.append style={fill=magenta}]{}{8}{8}
\ganttbar[bar/.append style={fill=magenta}]{}{12}{12}
\end{ganttchart}
\vspace{-10pt}
\caption{Schedule for the proposed work, measured in quarters from the
  beginning of the project. The lead PI of each Specific Aim will be RR
  (red), BQ (blue), and YNY (green). Project management will be
  continuous (black), but we anticipate extra attention will be
  necessary to align objectives in quarters 4, 8, and 12 (magenta).}
\label{fig:schedule}
\end{wrapfigure}
\textbf{Project management}: 
%
The success of the proposed research requires complementary expertise
and collaborative efforts in physics, applied mathematics, algorithms,
and computing. PI RR has been working on mathematical modeling with a
strong analytical background. PI YNY has been working on many areas of
computational fluid dynamics and applications to math biology for many
years. PI BQ has been working on integral equation methods, fast
algorithms, and their applications to fluid dynamics for many years.
Their recent collaborative work on the HAP model in two
dimensions~\cite{FuQuRyYo20} has provided a solid foundation for the
proposed research.

\smallskip

\textbf{Collaboration plan}: 
%
The management responsibility of this collaborative research will reside
with the lead PI (RR) for this endeavor. The research work is structured
to meet the tasks discussed in \S\ref{sec:proposed-work}. The PIs,
postdocs, and students will meet frequently on Zoom and in person when
possible. During the pandemic, the three PIs will meet weekly on Zoom
for a research meeting. Once the pandemic is under control, PI RR and PI
YNY will conduct biweekly meetings in person, with PI BQ Zoom in from
Florida. The PIs will share software packages, paper sources, and
references on a common \textsf{Git} repository. The resulting software
packages will be posted on the \textsf{Github} software repository.

\smallskip

\textbf{Research Schedule:} The detailed schedule for the proposed work
is shown in Figure~\ref{fig:schedule}. Each PI will spearhead a
different Specific Aim, and program management will happen
throughout, with special coordination taken in quarters 4, 8, and
12.

### 2021-11-13 NSF2021/Description/specificaim2.tex

When solving for the density functions $\eta_0$, $\eta_1$, and $\sigma$,
a carefully chosen IE formulation requires a mesh-independent number of
GMRES iterations~\cite{cam-ips-kel-mey-xue1996}. Therefore, the required
CPU time is proportional to the cost of a matrix-vector multiplication
that can be performed in optimal or near-optimal time with the fast
multipole method (FMM)~\cite{fmm5} and its extensions~\cite{fmm1, fmm2,
fmm3, fmm4, fmm6, fmm7, fmm8}. PI Quaife is experienced with applying
FMMs for the Stokes equation~\cite{qua-bir2014, bys-sha-qua2020} and the
screened Laplace equation~\cite{kro-qua2011, qua2011}. Furthermore, to
reduce the number of GMRES iterations, preconditioning strategies such
as the inverse Fast Multipole Method (IFMM)~\cite{cou-pou-dar2017} and
domain decomposition~\cite{che-bir2021}. PI Quaife used the IFMM to
precondition Stokes equations~\cite{qua-cou-dar2018}. A suite of other
preconditioners including sparse approximate inverses~\cite{che2000} and
multigrid~\cite{hem-sch1981, sch1982} will be investigated.

@@

A numerical issue when simulating the self-assembly of amphiphilic
particles is avoiding particle collision. The hydrophobic attraction
potential drives the amphiphilic particles towards one another and this
leads to physical contact in finite time. Such particle collisions in a
dense, rigid-body suspension are a great challenge and can be a
bottleneck in large-scale simulations. We propose to remedy this
computational challenge by developing novel high-order adaptive time
stepping methods. PI Quaife developed a high-order adaptive time
stepping method for hydrodynamic suspensions \cite{qua-bir2016}, and it
has served as a robust method to simulate processes including mixing and
adhesion in suspensions~\cite{qua-vee-you2019, kab-qua-bir2017}.
Following this previous work, the proposed work will use a spectral
deferred correction method~\cite{dut-gre-rok2000} since it iteratively
applies a low-order, single-step method to achieve high-order accuracy.
To estimate the error, at each time step, we will compute the total
force and torque of the system which is computationally cheap and
physically zero. PI Quaife's experience is that error estimates based on
physical constraints such as force- and torque-free conditions
appropriately determine how the time step should be adjusted so that the
dynamics are resolved without the computational expense of techniques
such as embedded Runge-Kutta methods, step-doubling, and Richardson
extrapolation.

>>>>

When solving for the density functions $\eta_0$, $\eta_1$, and $\sigma$,
a carefully chosen IE formulation requires a mesh-independent number of
GMRES iterations~\cite{cam-ips-kel-mey-xue1996}. Therefore, the required
CPU time is proportional to the cost of a matrix-vector multiplication
that can be performed in optimal or near-optimal time with the fast
multipole method (FMM)~\cite{fmm5} and its extensions~\cite{fmm1, fmm2,
fmm3, fmm4, fmm6, fmm7, fmm8}. PI BQ is experienced with applying FMMs
for the Stokes equation~\cite{qua-bir2014, bys-sha-qua2020} and the
screened Laplace equation~\cite{kro-qua2011, qua2011}. Furthermore, to
reduce the number of GMRES iterations, preconditioning strategies such
as the inverse Fast Multipole Method (IFMM)~\cite{cou-pou-dar2017} and
domain decomposition~\cite{che-bir2021}. PI BQ used the IFMM to
precondition Stokes equations~\cite{qua-cou-dar2018}. A suite of other
preconditioners including sparse approximate inverses~\cite{che2000} and
multigrid~\cite{hem-sch1981, sch1982} will be investigated.

@@

A numerical issue when simulating the self-assembly of amphiphilic
particles is avoiding particle collision. The hydrophobic attraction
potential drives the amphiphilic particles towards one another and this
leads to physical contact in finite time. Such particle collisions in a
dense, rigid-body suspension are a great challenge and can be a
bottleneck in large-scale simulations. We propose to remedy this
computational challenge by developing novel high-order adaptive time
stepping methods. PI BQ developed a high-order adaptive time
stepping method for hydrodynamic suspensions \cite{qua-bir2016}, and it
has served as a robust method to simulate processes including mixing and
adhesion in suspensions~\cite{qua-vee-you2019, kab-qua-bir2017}.
Following this previous work, the proposed work will use a spectral
deferred correction method~\cite{dut-gre-rok2000} since it iteratively
applies a low-order, single-step method to achieve high-order accuracy.
To estimate the error, at each time step, we will compute the total
force and torque of the system which is computationally cheap and
physically zero. PI BQ's experience is that error estimates based on
physical constraints such as force- and torque-free conditions
appropriately determine how the time step should be adjusted so that the
dynamics are resolved without the computational expense of techniques
such as embedded Runge-Kutta methods, step-doubling, and Richardson
extrapolation.
